---
title: "covid 19"
author: "Salena Bennett"
date: " Spring 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$: Covid 19 deaths in Korea Jan.-May are unrelated to sex.
$H_A$: Yes, there is a relation to Covid 19 deaths in Korea from Jan.- May relating to sex of the patient.

Table of patients:
```{r}
kRdata$sex = as.factor(kRdata$sex)
kRdata$result = as.factor(kRdata$result)                       
# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

`
```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
 Gender plays a role in the deaths of Covid 19 in Korea. Graph shows more males have died of Covid 19 than females with Covid 19 in Korea.





  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
Results shows Males died twice as much of Covid 19 than Females with Covid 19 in Korea. Showing that gender is playing a role in fatalities.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

P-value is so small we reject the null hypothesis. The probability of a male dying is 2.3 greater than of a Female. 


## Conclusions

The Conclusion is that males are more likely to die from Covid in Korea than Females that have Covid 19 in Korea. That Sex is a factor in the results of deaths of Covid 19 in Korea. 
